### 2021-05-05 README.Rmd
Flush out old files and start anew

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Code for Look-Ahead Screening Rules Paper

<!-- badges: start -->
[![R-CMD-check](https://github.com/jolars/LookAheadScreening/workflows/R-CMD-check/badge.svg)](https://github.com/jolars/LookAheadScreening/actions)
<!-- badges: end -->