### 2021-05-05 paper/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% Template for EYSM 2021 submissions %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please save the file as your name.
% Please, do not change any definitions.

\documentclass[b5paper,10pt,abstractoff,DIV=calc,headings=normal,twoside]{scrartcl}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,amsfonts,graphicx,etoolbox,scrlayer-scrpage}
\usepackage[noblocks]{authblk}
\makeatletter
\patchcmd{\@maketitle}{\huge}{\Large}{}{}
\patchcmd{\abstract}{\quotation}{}{}{}
\AtBeginEnvironment{abstract}{\noindent\ignorespaces}
\AtEndEnvironment{abstract}{\par\mbox{}}
\newcommand{\shortauthor}{}
\newcommand{\shorttitle}{\@title}
\makeatother
\setkomafont{author}{\small}
\setkomafont{title}{\rmfamily\bfseries}
\setkomafont{disposition}{\rmfamily\bfseries}
\renewcommand\Authfont{\bfseries}\renewcommand\Affilfont{\mdseries\slshape}
\def\AMS#1{\par\noindent \textbf{AMS subject classification: }#1\par}
\newcommand{\acknowledgements}{\par\mbox{}\par\noindent\textbf{Acknowledgements: }}
\newcommand{\keywords}[1]{\par\noindent\textbf{Keywords: }#1}
% \newcommand{\AMS}[1]{\par\noindent\textbf{AMS subject classification: }#1}
%\newcommand{\AMS}[1]{}
%\subject{\vspace{-2\baselineskip}}\date{\vspace{-2\baselineskip%}}
\renewcommand*{\titlepagestyle}{plain.scrheadings}\pagestyle{scrheadings}
\rehead[]{\shortauthor}\lohead[]{\shorttitle}\lehead[]{PPP}\rohead[]{PPP}
\cfoot[]{}\ofoot[]{}\ifoot[]{}\recalctypearea
\interfootnotelinepenalty=10000
%\usepackage{lipsum} %% generates dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% Available theorem environments %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{remark}
\newtheorem*{note}{Note}
\newtheorem{remark}{Remark}

\renewenvironment{abstract}{\bigskip\noindent\begin{minipage}{\textwidth}\setlength{\parindent}{15pt}\paragraph{Abstract:}}{\end{minipage}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Do not modify prior to this point %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{mathtools}
\usepackage[colorlinks]{hyperref}
\usepackage{url}
\usepackage[numbers]{natbib}
\bibliographystyle{abbrvnat}

% math operators
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% Please make sure to leave space for page numbers             %%%%%%%%%
%%%%%%%%% There is a placeholder PPP generated in the document to help %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\shortauthor}{J. Larsson}
\renewcommand{\shorttitle}{Look-Ahead Screening Rules for the Lasso}

\title{Look-Ahead Screening Rules for the Lasso}

\author[1]{Johan Larsson%
  \thanks{%
    \href{mailto:johan.larsson@stat.lu.se}{\url{johan.larsson@stat.lu.se}}
  }%
}
\affil[1]{The Department of Statistics, Lund University}

\maketitle

\begin{abstract}
  The lasso is a popular method for inducing shrinkage and sparsity in the
  solution vector (coefficients) of regression problems, particularly when
  the number of predictors far outnumber the number of observations. Solving
  the lasso for high-dimensional data can, however, be computationally
  demanding. Fortunately, this computational load can be alleviated via the
  use of \emph{screening rules}, which screen and discard predictors prior to
  fitting the model, leading a reduced problem to be solved. Screening rules
  are particularly effective when fitting a full regularization path: a
  sequence of models with decreasing penalization. Screening rules can be
  safe or heuristic. Safe rules certify that discarded predictors are not in
  the solution; heuristic ones do not. Existing screening rules typically
  work sequentially or dynamically. Sequential rules screen predictors for
  the next model along the regularization path, whereas dynamical rules
  screen during optimization of the current model. There has, however,
  previously been no attempts to design screening rules that screen further
  along the path.

  In this paper, we present a new screening strategy: \emph{look-ahead}
  screening rules. Our method uses safe screening rules to find a range of
  penalty values for which a given predictor cannot enter the model, thereby
  screening predictors along the remainder of the path. Our screening rules
  lead to reductions in the time required for screening and also applies to
  heuristic rules, for which the time required to conduct checks of the
  optimality conditions to guard against violations of the rules is reduced.
  In experiments we show that these look-ahead screening rules improve the
  performance of existing screening strategies and that the additional cost
  of screening ahead on the path is marginal.
\end{abstract}

\keywords{lasso, sparse regression, screening rules, safe screening rules}

\smallskip

\AMS{62J07}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

\section{Preliminaries}%
\label{sec:preliminaries}

Starting with the preliminaries of our problem, we take \(X \in \mathbb{R}^{n
  \times p}\) be the design matrix of \(n\) observations and \(p\) predictors
and \(y \in \mathbb{R}^n\) the response vector.

The lasso is represented by the following convex optimization problem:
\begin{equation}
  \label{eq:primal}
  \operatorname*{minimize}_{\beta \in \mathbb{R}^p} P(\beta; \lambda)
\end{equation}
where
\begin{equation}
  P(\beta; \lambda) =
  \frac 1 2 \lVert y - X\beta \rVert_2^2 + \lambda \lVert \beta \rVert_1
\end{equation}
is the \emph{primal} objective. We let \(\hat \beta_\lambda\) be the solution to
\eqref{eq:primal} for a given \(\lambda\). In addition, we also let \(\tilde
\beta_\lambda\) be a numerical approximation to \(\hat\beta_\lambda\)
returned by an algorithm
tasked with solving \eqref{eq:primal}.

The Karush--Kuhn--Tucker (KKT) stationarity condition for \eqref{eq:primal}
holds that
\begin{equation}
  \label{eq:stationarity}
  0 \in X^T(X\beta - y) + \lambda\partial,
\end{equation}
where \(\partial\) is the subdifferential of the \(\ell_1\)-norm, with
its \(j\)th element corresponding to
\begin{equation}
  \label{eq:subdifferential}
  \partial_j \in
  \begin{cases}
    \{\sign(\beta_j)\} & \text{if } \beta_j \neq 0 \\
    [-1,1]             & \text{otherwise.}
  \end{cases}
\end{equation}
Moreover, we define the dual problem of \eqref{eq:primal} as
\begin{equation}
  \label{eq:dual}
  \operatorname*{maximize}_{\theta \in \mathbb{R}^n} D(\theta; \lambda)
\end{equation}
where
\begin{equation}
  D(\theta) = \frac 12 y^T y - \frac{\lambda^2}{2} \left\lVert \theta
  - \frac y \lambda \right\rVert_2^2
\end{equation}
is the \emph{dual} objective.
The relationship between the primal and dual problems is given by
\[
  y = X\hat\beta_\lambda + \lambda \hat\theta_\lambda.
\]
Next, we let \(G\) be the so-called \emph{duality gap}, which we define
as
\begin{equation}
  \begin{aligned}
    G(\beta, \theta; \lambda)
     & = P(\beta; \lambda) - D(\theta; \lambda)                                                 \\
     & = \frac 12 \lVert y - X\beta\rVert_2^2 + \lambda \lVert \beta \rVert_1
    - \frac 12 y^T y + \frac{\lambda^2}{2} \left\lVert \theta - \frac y \theta \right\rVert_2^2 \\
     & =
    \frac 12 \lVert y - X\beta\rVert_2^2 + \lambda \lVert \beta \rVert_1
    - \lambda \theta^T y + \frac{\lambda^2}{2} \theta^T \theta.
  \end{aligned}
\end{equation}
In the case of the lasso, strong duality holds, which means that
\(G(\tilde\beta_\lambda, \tilde\theta_\lambda; \lambda) = 0\) for any
choice of \(\lambda\).

Suppose, now, that we have solved the lasso for \(\lambda\); then
for any given \(\lambda^* \geq \lambda\),
the Gap Safe rule~\citep{ndiaye2017} for the lasso discards the \(j\)th
predictor if
\begin{equation}
  \label{eq:gap-safe-rule}
  |X^T \tilde \theta_\lambda|_j + \lVert x_j\rVert_2
  \sqrt{\frac{1}{\lambda_*^2}
    G(\tilde \beta_\lambda, \tilde \theta_\lambda; \lambda^*)}
  < 1
\end{equation}
where
\[
  \tilde\theta_\lambda = \frac{y - X\tilde\beta_\lambda}{
    \max\big( \max_j|x_j^T(y - X\tilde\beta_\lambda)|, \lambda\big)}
\]
is a dual-feasible point~\cite{ndiaye2017}, which we can always obtain
by dual scaling whenever a candidate is infeasible. Using these facts,
we now arrive at \autoref{thm:look-ahead}, which is the main result of
this paper.

\begin{theorem}
  \label{thm:look-ahead}
  Let
  \((\tilde\beta_\lambda,\tilde\theta_\lambda)\) be
  a feasible primal--dual point for the solution to
  \eqref{eq:primal} and \eqref{eq:dual}. Then
  \(\big(\hat\beta_{\lambda^*}\big)_j = 0\) for any
  \[
    \lambda_* > \frac{-b + \sqrt{b^2 - 4ac}}{2a}
  \]
  where
  \[
    \begin{aligned}
      a & = \big( 1 - | x_j^T \tilde\theta_\lambda|\big)^2 -
      \frac 12 \tilde\theta_\lambda^T \tilde\theta_\lambda \lVert x_j\rVert_2^2      \\
      b & = \big(\tilde\theta_\lambda^T y - \lVert \tilde\beta_\lambda \rVert_1\big)
      \lVert x_j \rVert_2^2                                                          \\
      c & = - \frac 12 \lVert y - X\tilde\beta_\lambda\rVert_2^2
      \lVert x_j\rVert_2^2.                                                          \\
    \end{aligned}
  \]
\end{theorem}
\begin{proof}

\end{proof}

Note that there must be some \(\lambda_*^2\), for which it holds that
\[
  |X^T \tilde \theta_\lambda|_j +
  \lVert x_j\rVert_2 \sqrt{\frac{1}{\lambda_*^2} G(\tilde \beta_\lambda, \tilde \theta_\lambda; \lambda^*)}
  = 1
\]

\acknowledgements{Your acknowledgements.}

\bibliography{LookAheadScreening.bib}

\end{document}