### 2021-05-12 paper/figures/casestudy.tex
Update results

% Created by tikzDevice version 0.12.3.1 on 2021-05-11 11:26:34
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (339.67,289.08);
\begin{scope}
\path[clip] ( 25.95,117.46) rectangle (335.17,192.22);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 25.95,129.36) --
	(335.17,129.36);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 25.95,144.65) --
	(335.17,144.65);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 25.95,161.64) --
	(335.17,161.64);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 25.95,178.63) --
	(335.17,178.63);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 46.33,117.46) --
	( 46.33,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 78.62,117.46) --
	( 78.62,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (112.60,117.46) --
	(112.60,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (146.58,117.46) --
	(146.58,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (180.56,117.46) --
	(180.56,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (214.54,117.46) --
	(214.54,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (248.52,117.46) --
	(248.52,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (282.50,117.46) --
	(282.50,192.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] (316.48,117.46) --
	(316.48,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 25.95,122.56) --
	(335.17,122.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 25.95,136.15) --
	(335.17,136.15);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 25.95,153.14) --
	(335.17,153.14);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 25.95,170.13) --
	(335.17,170.13);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 25.95,187.12) --
	(335.17,187.12);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 31.04,117.46) --
	( 31.04,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 61.63,117.46) --
	( 61.63,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 95.61,117.46) --
	( 95.61,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (129.59,117.46) --
	(129.59,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (163.57,117.46) --
	(163.57,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (197.55,117.46) --
	(197.55,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (231.53,117.46) --
	(231.53,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (265.51,117.46) --
	(265.51,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (299.49,117.46) --
	(299.49,192.22);

\path[draw=drawColor,line width= 0.5pt,line join=round] (333.47,117.46) --
	(333.47,192.22);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,120.86) rectangle ( 32.74,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,120.86) rectangle ( 36.14,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,120.86) rectangle ( 39.54,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,120.86) rectangle ( 42.94,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,120.86) rectangle ( 46.33,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,120.86) rectangle ( 49.73,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,120.86) rectangle ( 53.13,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,120.86) rectangle ( 56.53,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,120.86) rectangle ( 59.93,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,120.86) rectangle ( 63.32,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,120.86) rectangle ( 66.72,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,120.86) rectangle ( 70.12,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,120.86) rectangle ( 73.52,124.26);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,120.86) rectangle ( 76.92,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,120.86) rectangle ( 80.32,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,120.86) rectangle ( 83.71,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,120.86) rectangle ( 87.11,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,120.86) rectangle ( 90.51,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,120.86) rectangle ( 93.91,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,120.86) rectangle ( 97.31,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,120.86) rectangle (100.70,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,120.86) rectangle (104.10,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,120.86) rectangle (107.50,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,120.86) rectangle (110.90,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,120.86) rectangle (114.30,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,120.86) rectangle (117.69,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,120.86) rectangle (121.09,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,120.86) rectangle (124.49,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,120.86) rectangle (127.89,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,120.86) rectangle (131.29,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,120.86) rectangle (134.68,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,120.86) rectangle (138.08,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,120.86) rectangle (141.48,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,120.86) rectangle (144.88,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,120.86) rectangle (148.28,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,120.86) rectangle (151.67,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,120.86) rectangle (155.07,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,120.86) rectangle (158.47,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,120.86) rectangle (161.87,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,120.86) rectangle (165.27,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,120.86) rectangle (168.66,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,120.86) rectangle (172.06,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,120.86) rectangle (175.46,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,120.86) rectangle (178.86,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,120.86) rectangle (182.26,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,120.86) rectangle (185.65,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,120.86) rectangle (189.05,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,120.86) rectangle (192.45,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,120.86) rectangle (195.85,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,120.86) rectangle (199.25,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,120.86) rectangle (202.64,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,120.86) rectangle (206.04,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,120.86) rectangle (209.44,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,120.86) rectangle (212.84,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,120.86) rectangle (216.24,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,120.86) rectangle (219.64,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,120.86) rectangle (223.03,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,120.86) rectangle (226.43,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,120.86) rectangle (229.83,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,120.86) rectangle (233.23,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,120.86) rectangle (236.63,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,120.86) rectangle (240.02,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,120.86) rectangle (243.42,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,120.86) rectangle (246.82,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,120.86) rectangle (250.22,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,120.86) rectangle (253.62,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,120.86) rectangle (257.01,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,120.86) rectangle (260.41,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,120.86) rectangle (263.81,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,120.86) rectangle (267.21,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,120.86) rectangle (270.61,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,120.86) rectangle (274.00,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,120.86) rectangle (277.40,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,120.86) rectangle (280.80,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,120.86) rectangle (284.20,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,120.86) rectangle (287.60,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,120.86) rectangle (290.99,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,120.86) rectangle (294.39,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,120.86) rectangle (297.79,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,120.86) rectangle (301.19,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,120.86) rectangle (304.59,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,120.86) rectangle (307.98,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,120.86) rectangle (311.38,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,120.86) rectangle (314.78,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,120.86) rectangle (318.18,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,120.86) rectangle (321.58,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,120.86) rectangle (324.97,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,120.86) rectangle (328.37,124.26);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,120.86) rectangle (331.77,124.26);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,124.26) rectangle ( 32.74,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,124.26) rectangle ( 36.14,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,124.26) rectangle ( 39.54,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,124.26) rectangle ( 42.94,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,124.26) rectangle ( 46.33,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,124.26) rectangle ( 49.73,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,124.26) rectangle ( 53.13,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,124.26) rectangle ( 56.53,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,124.26) rectangle ( 59.93,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,124.26) rectangle ( 63.32,127.66);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,124.26) rectangle ( 66.72,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,124.26) rectangle ( 70.12,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,124.26) rectangle ( 73.52,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,124.26) rectangle ( 76.92,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,124.26) rectangle ( 80.32,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,124.26) rectangle ( 83.71,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,124.26) rectangle ( 87.11,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,124.26) rectangle ( 90.51,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,124.26) rectangle ( 93.91,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,124.26) rectangle ( 97.31,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,124.26) rectangle (100.70,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,124.26) rectangle (104.10,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,124.26) rectangle (107.50,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,124.26) rectangle (110.90,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,124.26) rectangle (114.30,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,124.26) rectangle (117.69,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,124.26) rectangle (121.09,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,124.26) rectangle (124.49,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,124.26) rectangle (127.89,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,124.26) rectangle (131.29,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,124.26) rectangle (134.68,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,124.26) rectangle (138.08,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,124.26) rectangle (141.48,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,124.26) rectangle (144.88,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,124.26) rectangle (148.28,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,124.26) rectangle (151.67,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,124.26) rectangle (155.07,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,124.26) rectangle (158.47,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,124.26) rectangle (161.87,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,124.26) rectangle (165.27,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,124.26) rectangle (168.66,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,124.26) rectangle (172.06,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,124.26) rectangle (175.46,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,124.26) rectangle (178.86,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,124.26) rectangle (182.26,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,124.26) rectangle (185.65,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,124.26) rectangle (189.05,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,124.26) rectangle (192.45,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,124.26) rectangle (195.85,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,124.26) rectangle (199.25,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,124.26) rectangle (202.64,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,124.26) rectangle (206.04,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,124.26) rectangle (209.44,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,124.26) rectangle (212.84,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,124.26) rectangle (216.24,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,124.26) rectangle (219.64,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,124.26) rectangle (223.03,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,124.26) rectangle (226.43,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,124.26) rectangle (229.83,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,124.26) rectangle (233.23,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,124.26) rectangle (236.63,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,124.26) rectangle (240.02,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,124.26) rectangle (243.42,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,124.26) rectangle (246.82,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,124.26) rectangle (250.22,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,124.26) rectangle (253.62,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,124.26) rectangle (257.01,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,124.26) rectangle (260.41,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,124.26) rectangle (263.81,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,124.26) rectangle (267.21,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,124.26) rectangle (270.61,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,124.26) rectangle (274.00,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,124.26) rectangle (277.40,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,124.26) rectangle (280.80,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,124.26) rectangle (284.20,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,124.26) rectangle (287.60,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,124.26) rectangle (290.99,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,124.26) rectangle (294.39,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,124.26) rectangle (297.79,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,124.26) rectangle (301.19,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,124.26) rectangle (304.59,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,124.26) rectangle (307.98,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,124.26) rectangle (311.38,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,124.26) rectangle (314.78,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,124.26) rectangle (318.18,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,124.26) rectangle (321.58,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,124.26) rectangle (324.97,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,124.26) rectangle (328.37,127.66);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,124.26) rectangle (331.77,127.66);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,127.66) rectangle ( 32.74,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,127.66) rectangle ( 36.14,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,127.66) rectangle ( 39.54,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,127.66) rectangle ( 42.94,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,127.66) rectangle ( 46.33,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,127.66) rectangle ( 49.73,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,127.66) rectangle ( 53.13,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,127.66) rectangle ( 56.53,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,127.66) rectangle ( 59.93,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,127.66) rectangle ( 63.32,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,127.66) rectangle ( 66.72,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,127.66) rectangle ( 70.12,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,127.66) rectangle ( 73.52,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,127.66) rectangle ( 76.92,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,127.66) rectangle ( 80.32,131.06);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,127.66) rectangle ( 83.71,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,127.66) rectangle ( 87.11,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,127.66) rectangle ( 90.51,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,127.66) rectangle ( 93.91,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,127.66) rectangle ( 97.31,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,127.66) rectangle (100.70,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,127.66) rectangle (104.10,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,127.66) rectangle (107.50,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,127.66) rectangle (110.90,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,127.66) rectangle (114.30,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,127.66) rectangle (117.69,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,127.66) rectangle (121.09,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,127.66) rectangle (124.49,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,127.66) rectangle (127.89,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,127.66) rectangle (131.29,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,127.66) rectangle (134.68,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,127.66) rectangle (138.08,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,127.66) rectangle (141.48,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,127.66) rectangle (144.88,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,127.66) rectangle (148.28,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,127.66) rectangle (151.67,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,127.66) rectangle (155.07,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,127.66) rectangle (158.47,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,127.66) rectangle (161.87,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,127.66) rectangle (165.27,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,127.66) rectangle (168.66,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,127.66) rectangle (172.06,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,127.66) rectangle (175.46,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,127.66) rectangle (178.86,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,127.66) rectangle (182.26,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,127.66) rectangle (185.65,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,127.66) rectangle (189.05,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,127.66) rectangle (192.45,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,127.66) rectangle (195.85,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,127.66) rectangle (199.25,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,127.66) rectangle (202.64,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,127.66) rectangle (206.04,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,127.66) rectangle (209.44,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,127.66) rectangle (212.84,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,127.66) rectangle (216.24,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,127.66) rectangle (219.64,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,127.66) rectangle (223.03,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,127.66) rectangle (226.43,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,127.66) rectangle (229.83,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,127.66) rectangle (233.23,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,127.66) rectangle (236.63,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,127.66) rectangle (240.02,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,127.66) rectangle (243.42,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,127.66) rectangle (246.82,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,127.66) rectangle (250.22,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,127.66) rectangle (253.62,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,127.66) rectangle (257.01,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,127.66) rectangle (260.41,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,127.66) rectangle (263.81,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,127.66) rectangle (267.21,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,127.66) rectangle (270.61,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,127.66) rectangle (274.00,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,127.66) rectangle (277.40,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,127.66) rectangle (280.80,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,127.66) rectangle (284.20,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,127.66) rectangle (287.60,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,127.66) rectangle (290.99,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,127.66) rectangle (294.39,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,127.66) rectangle (297.79,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,127.66) rectangle (301.19,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,127.66) rectangle (304.59,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,127.66) rectangle (307.98,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,127.66) rectangle (311.38,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,127.66) rectangle (314.78,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,127.66) rectangle (318.18,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,127.66) rectangle (321.58,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,127.66) rectangle (324.97,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,127.66) rectangle (328.37,131.06);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,127.66) rectangle (331.77,131.06);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,131.06) rectangle ( 32.74,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,131.06) rectangle ( 36.14,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,131.06) rectangle ( 39.54,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,131.06) rectangle ( 42.94,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,131.06) rectangle ( 46.33,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,131.06) rectangle ( 49.73,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,131.06) rectangle ( 53.13,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,131.06) rectangle ( 56.53,134.45);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,131.06) rectangle ( 59.93,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,131.06) rectangle ( 63.32,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,131.06) rectangle ( 66.72,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,131.06) rectangle ( 70.12,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,131.06) rectangle ( 73.52,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,131.06) rectangle ( 76.92,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,131.06) rectangle ( 80.32,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,131.06) rectangle ( 83.71,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,131.06) rectangle ( 87.11,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,131.06) rectangle ( 90.51,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,131.06) rectangle ( 93.91,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,131.06) rectangle ( 97.31,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,131.06) rectangle (100.70,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,131.06) rectangle (104.10,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,131.06) rectangle (107.50,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,131.06) rectangle (110.90,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,131.06) rectangle (114.30,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,131.06) rectangle (117.69,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,131.06) rectangle (121.09,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,131.06) rectangle (124.49,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,131.06) rectangle (127.89,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,131.06) rectangle (131.29,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,131.06) rectangle (134.68,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,131.06) rectangle (138.08,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,131.06) rectangle (141.48,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,131.06) rectangle (144.88,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,131.06) rectangle (148.28,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,131.06) rectangle (151.67,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,131.06) rectangle (155.07,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,131.06) rectangle (158.47,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,131.06) rectangle (161.87,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,131.06) rectangle (165.27,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,131.06) rectangle (168.66,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,131.06) rectangle (172.06,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,131.06) rectangle (175.46,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,131.06) rectangle (178.86,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,131.06) rectangle (182.26,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,131.06) rectangle (185.65,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,131.06) rectangle (189.05,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,131.06) rectangle (192.45,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,131.06) rectangle (195.85,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,131.06) rectangle (199.25,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,131.06) rectangle (202.64,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,131.06) rectangle (206.04,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,131.06) rectangle (209.44,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,131.06) rectangle (212.84,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,131.06) rectangle (216.24,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,131.06) rectangle (219.64,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,131.06) rectangle (223.03,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,131.06) rectangle (226.43,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,131.06) rectangle (229.83,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,131.06) rectangle (233.23,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,131.06) rectangle (236.63,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,131.06) rectangle (240.02,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,131.06) rectangle (243.42,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,131.06) rectangle (246.82,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,131.06) rectangle (250.22,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,131.06) rectangle (253.62,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,131.06) rectangle (257.01,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,131.06) rectangle (260.41,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,131.06) rectangle (263.81,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,131.06) rectangle (267.21,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,131.06) rectangle (270.61,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,131.06) rectangle (274.00,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,131.06) rectangle (277.40,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,131.06) rectangle (280.80,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,131.06) rectangle (284.20,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,131.06) rectangle (287.60,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,131.06) rectangle (290.99,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,131.06) rectangle (294.39,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,131.06) rectangle (297.79,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,131.06) rectangle (301.19,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,131.06) rectangle (304.59,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,131.06) rectangle (307.98,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,131.06) rectangle (311.38,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,131.06) rectangle (314.78,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,131.06) rectangle (318.18,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,131.06) rectangle (321.58,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,131.06) rectangle (324.97,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,131.06) rectangle (328.37,134.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,131.06) rectangle (331.77,134.45);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,134.45) rectangle ( 32.74,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,134.45) rectangle ( 36.14,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,134.45) rectangle ( 39.54,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,134.45) rectangle ( 42.94,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,134.45) rectangle ( 46.33,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,134.45) rectangle ( 49.73,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,134.45) rectangle ( 53.13,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,134.45) rectangle ( 56.53,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,134.45) rectangle ( 59.93,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,134.45) rectangle ( 63.32,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,134.45) rectangle ( 66.72,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,134.45) rectangle ( 70.12,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,134.45) rectangle ( 73.52,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,134.45) rectangle ( 76.92,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,134.45) rectangle ( 80.32,137.85);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,134.45) rectangle ( 83.71,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,134.45) rectangle ( 87.11,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,134.45) rectangle ( 90.51,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,134.45) rectangle ( 93.91,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,134.45) rectangle ( 97.31,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,134.45) rectangle (100.70,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,134.45) rectangle (104.10,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,134.45) rectangle (107.50,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,134.45) rectangle (110.90,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,134.45) rectangle (114.30,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,134.45) rectangle (117.69,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,134.45) rectangle (121.09,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,134.45) rectangle (124.49,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,134.45) rectangle (127.89,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,134.45) rectangle (131.29,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,134.45) rectangle (134.68,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,134.45) rectangle (138.08,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,134.45) rectangle (141.48,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,134.45) rectangle (144.88,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,134.45) rectangle (148.28,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,134.45) rectangle (151.67,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,134.45) rectangle (155.07,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,134.45) rectangle (158.47,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,134.45) rectangle (161.87,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,134.45) rectangle (165.27,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,134.45) rectangle (168.66,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,134.45) rectangle (172.06,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,134.45) rectangle (175.46,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,134.45) rectangle (178.86,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,134.45) rectangle (182.26,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,134.45) rectangle (185.65,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,134.45) rectangle (189.05,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,134.45) rectangle (192.45,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,134.45) rectangle (195.85,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,134.45) rectangle (199.25,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,134.45) rectangle (202.64,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,134.45) rectangle (206.04,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,134.45) rectangle (209.44,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,134.45) rectangle (212.84,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,134.45) rectangle (216.24,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,134.45) rectangle (219.64,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,134.45) rectangle (223.03,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,134.45) rectangle (226.43,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,134.45) rectangle (229.83,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,134.45) rectangle (233.23,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,134.45) rectangle (236.63,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,134.45) rectangle (240.02,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,134.45) rectangle (243.42,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,134.45) rectangle (246.82,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,134.45) rectangle (250.22,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,134.45) rectangle (253.62,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,134.45) rectangle (257.01,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,134.45) rectangle (260.41,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,134.45) rectangle (263.81,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,134.45) rectangle (267.21,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,134.45) rectangle (270.61,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,134.45) rectangle (274.00,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,134.45) rectangle (277.40,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,134.45) rectangle (280.80,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,134.45) rectangle (284.20,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,134.45) rectangle (287.60,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,134.45) rectangle (290.99,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,134.45) rectangle (294.39,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,134.45) rectangle (297.79,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,134.45) rectangle (301.19,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,134.45) rectangle (304.59,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,134.45) rectangle (307.98,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,134.45) rectangle (311.38,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,134.45) rectangle (314.78,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,134.45) rectangle (318.18,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,134.45) rectangle (321.58,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,134.45) rectangle (324.97,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,134.45) rectangle (328.37,137.85);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,134.45) rectangle (331.77,137.85);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,137.85) rectangle ( 32.74,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,137.85) rectangle ( 36.14,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,137.85) rectangle ( 39.54,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,137.85) rectangle ( 42.94,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,137.85) rectangle ( 46.33,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,137.85) rectangle ( 49.73,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,137.85) rectangle ( 53.13,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,137.85) rectangle ( 56.53,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,137.85) rectangle ( 59.93,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,137.85) rectangle ( 63.32,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,137.85) rectangle ( 66.72,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,137.85) rectangle ( 70.12,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,137.85) rectangle ( 73.52,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,137.85) rectangle ( 76.92,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,137.85) rectangle ( 80.32,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,137.85) rectangle ( 83.71,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,137.85) rectangle ( 87.11,141.25);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,137.85) rectangle ( 90.51,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,137.85) rectangle ( 93.91,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,137.85) rectangle ( 97.31,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,137.85) rectangle (100.70,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,137.85) rectangle (104.10,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,137.85) rectangle (107.50,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,137.85) rectangle (110.90,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,137.85) rectangle (114.30,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,137.85) rectangle (117.69,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,137.85) rectangle (121.09,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,137.85) rectangle (124.49,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,137.85) rectangle (127.89,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,137.85) rectangle (131.29,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,137.85) rectangle (134.68,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,137.85) rectangle (138.08,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,137.85) rectangle (141.48,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,137.85) rectangle (144.88,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,137.85) rectangle (148.28,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,137.85) rectangle (151.67,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,137.85) rectangle (155.07,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,137.85) rectangle (158.47,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,137.85) rectangle (161.87,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,137.85) rectangle (165.27,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,137.85) rectangle (168.66,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,137.85) rectangle (172.06,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,137.85) rectangle (175.46,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,137.85) rectangle (178.86,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,137.85) rectangle (182.26,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,137.85) rectangle (185.65,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,137.85) rectangle (189.05,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,137.85) rectangle (192.45,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,137.85) rectangle (195.85,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,137.85) rectangle (199.25,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,137.85) rectangle (202.64,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,137.85) rectangle (206.04,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,137.85) rectangle (209.44,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,137.85) rectangle (212.84,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,137.85) rectangle (216.24,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,137.85) rectangle (219.64,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,137.85) rectangle (223.03,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,137.85) rectangle (226.43,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,137.85) rectangle (229.83,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,137.85) rectangle (233.23,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,137.85) rectangle (236.63,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,137.85) rectangle (240.02,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,137.85) rectangle (243.42,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,137.85) rectangle (246.82,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,137.85) rectangle (250.22,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,137.85) rectangle (253.62,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,137.85) rectangle (257.01,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,137.85) rectangle (260.41,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,137.85) rectangle (263.81,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,137.85) rectangle (267.21,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,137.85) rectangle (270.61,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,137.85) rectangle (274.00,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,137.85) rectangle (277.40,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,137.85) rectangle (280.80,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,137.85) rectangle (284.20,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,137.85) rectangle (287.60,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,137.85) rectangle (290.99,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,137.85) rectangle (294.39,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,137.85) rectangle (297.79,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,137.85) rectangle (301.19,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,137.85) rectangle (304.59,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,137.85) rectangle (307.98,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,137.85) rectangle (311.38,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,137.85) rectangle (314.78,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,137.85) rectangle (318.18,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,137.85) rectangle (321.58,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,137.85) rectangle (324.97,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,137.85) rectangle (328.37,141.25);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,137.85) rectangle (331.77,141.25);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,141.25) rectangle ( 32.74,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,141.25) rectangle ( 36.14,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,141.25) rectangle ( 39.54,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,141.25) rectangle ( 42.94,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,141.25) rectangle ( 46.33,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,141.25) rectangle ( 49.73,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,141.25) rectangle ( 53.13,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,141.25) rectangle ( 56.53,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,141.25) rectangle ( 59.93,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,141.25) rectangle ( 63.32,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,141.25) rectangle ( 66.72,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,141.25) rectangle ( 70.12,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,141.25) rectangle ( 73.52,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,141.25) rectangle ( 76.92,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,141.25) rectangle ( 80.32,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,141.25) rectangle ( 83.71,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,141.25) rectangle ( 87.11,144.65);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,141.25) rectangle ( 90.51,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,141.25) rectangle ( 93.91,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,141.25) rectangle ( 97.31,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,141.25) rectangle (100.70,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,141.25) rectangle (104.10,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,141.25) rectangle (107.50,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,141.25) rectangle (110.90,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,141.25) rectangle (114.30,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,141.25) rectangle (117.69,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,141.25) rectangle (121.09,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,141.25) rectangle (124.49,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,141.25) rectangle (127.89,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,141.25) rectangle (131.29,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,141.25) rectangle (134.68,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,141.25) rectangle (138.08,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,141.25) rectangle (141.48,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,141.25) rectangle (144.88,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,141.25) rectangle (148.28,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,141.25) rectangle (151.67,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,141.25) rectangle (155.07,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,141.25) rectangle (158.47,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,141.25) rectangle (161.87,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,141.25) rectangle (165.27,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,141.25) rectangle (168.66,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,141.25) rectangle (172.06,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,141.25) rectangle (175.46,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,141.25) rectangle (178.86,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,141.25) rectangle (182.26,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,141.25) rectangle (185.65,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,141.25) rectangle (189.05,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,141.25) rectangle (192.45,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,141.25) rectangle (195.85,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,141.25) rectangle (199.25,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,141.25) rectangle (202.64,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,141.25) rectangle (206.04,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,141.25) rectangle (209.44,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,141.25) rectangle (212.84,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,141.25) rectangle (216.24,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,141.25) rectangle (219.64,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,141.25) rectangle (223.03,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,141.25) rectangle (226.43,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,141.25) rectangle (229.83,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,141.25) rectangle (233.23,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,141.25) rectangle (236.63,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,141.25) rectangle (240.02,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,141.25) rectangle (243.42,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,141.25) rectangle (246.82,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,141.25) rectangle (250.22,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,141.25) rectangle (253.62,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,141.25) rectangle (257.01,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,141.25) rectangle (260.41,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,141.25) rectangle (263.81,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,141.25) rectangle (267.21,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,141.25) rectangle (270.61,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,141.25) rectangle (274.00,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,141.25) rectangle (277.40,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,141.25) rectangle (280.80,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,141.25) rectangle (284.20,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,141.25) rectangle (287.60,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,141.25) rectangle (290.99,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,141.25) rectangle (294.39,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,141.25) rectangle (297.79,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,141.25) rectangle (301.19,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,141.25) rectangle (304.59,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,141.25) rectangle (307.98,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,141.25) rectangle (311.38,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,141.25) rectangle (314.78,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,141.25) rectangle (318.18,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,141.25) rectangle (321.58,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,141.25) rectangle (324.97,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,141.25) rectangle (328.37,144.65);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,141.25) rectangle (331.77,144.65);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,144.65) rectangle ( 32.74,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,144.65) rectangle ( 36.14,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,144.65) rectangle ( 39.54,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,144.65) rectangle ( 42.94,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,144.65) rectangle ( 46.33,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,144.65) rectangle ( 49.73,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,144.65) rectangle ( 53.13,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,144.65) rectangle ( 56.53,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,144.65) rectangle ( 59.93,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,144.65) rectangle ( 63.32,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,144.65) rectangle ( 66.72,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,144.65) rectangle ( 70.12,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,144.65) rectangle ( 73.52,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,144.65) rectangle ( 76.92,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,144.65) rectangle ( 80.32,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,144.65) rectangle ( 83.71,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,144.65) rectangle ( 87.11,148.05);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,144.65) rectangle ( 90.51,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,144.65) rectangle ( 93.91,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,144.65) rectangle ( 97.31,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,144.65) rectangle (100.70,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,144.65) rectangle (104.10,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,144.65) rectangle (107.50,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,144.65) rectangle (110.90,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,144.65) rectangle (114.30,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,144.65) rectangle (117.69,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,144.65) rectangle (121.09,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,144.65) rectangle (124.49,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,144.65) rectangle (127.89,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,144.65) rectangle (131.29,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,144.65) rectangle (134.68,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,144.65) rectangle (138.08,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,144.65) rectangle (141.48,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,144.65) rectangle (144.88,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,144.65) rectangle (148.28,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,144.65) rectangle (151.67,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,144.65) rectangle (155.07,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,144.65) rectangle (158.47,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,144.65) rectangle (161.87,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,144.65) rectangle (165.27,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,144.65) rectangle (168.66,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,144.65) rectangle (172.06,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,144.65) rectangle (175.46,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,144.65) rectangle (178.86,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,144.65) rectangle (182.26,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,144.65) rectangle (185.65,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,144.65) rectangle (189.05,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,144.65) rectangle (192.45,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,144.65) rectangle (195.85,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,144.65) rectangle (199.25,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,144.65) rectangle (202.64,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,144.65) rectangle (206.04,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,144.65) rectangle (209.44,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,144.65) rectangle (212.84,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,144.65) rectangle (216.24,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,144.65) rectangle (219.64,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,144.65) rectangle (223.03,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,144.65) rectangle (226.43,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,144.65) rectangle (229.83,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,144.65) rectangle (233.23,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,144.65) rectangle (236.63,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,144.65) rectangle (240.02,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,144.65) rectangle (243.42,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,144.65) rectangle (246.82,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,144.65) rectangle (250.22,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,144.65) rectangle (253.62,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,144.65) rectangle (257.01,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,144.65) rectangle (260.41,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,144.65) rectangle (263.81,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,144.65) rectangle (267.21,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,144.65) rectangle (270.61,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,144.65) rectangle (274.00,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,144.65) rectangle (277.40,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,144.65) rectangle (280.80,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,144.65) rectangle (284.20,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,144.65) rectangle (287.60,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,144.65) rectangle (290.99,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,144.65) rectangle (294.39,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,144.65) rectangle (297.79,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,144.65) rectangle (301.19,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,144.65) rectangle (304.59,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,144.65) rectangle (307.98,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,144.65) rectangle (311.38,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,144.65) rectangle (314.78,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,144.65) rectangle (318.18,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,144.65) rectangle (321.58,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,144.65) rectangle (324.97,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,144.65) rectangle (328.37,148.05);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,144.65) rectangle (331.77,148.05);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,148.05) rectangle ( 32.74,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,148.05) rectangle ( 36.14,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,148.05) rectangle ( 39.54,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,148.05) rectangle ( 42.94,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,148.05) rectangle ( 46.33,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,148.05) rectangle ( 49.73,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,148.05) rectangle ( 53.13,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,148.05) rectangle ( 56.53,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,148.05) rectangle ( 59.93,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,148.05) rectangle ( 63.32,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,148.05) rectangle ( 66.72,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,148.05) rectangle ( 70.12,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,148.05) rectangle ( 73.52,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,148.05) rectangle ( 76.92,151.45);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,148.05) rectangle ( 80.32,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,148.05) rectangle ( 83.71,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,148.05) rectangle ( 87.11,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,148.05) rectangle ( 90.51,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,148.05) rectangle ( 93.91,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,148.05) rectangle ( 97.31,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,148.05) rectangle (100.70,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,148.05) rectangle (104.10,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,148.05) rectangle (107.50,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,148.05) rectangle (110.90,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,148.05) rectangle (114.30,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,148.05) rectangle (117.69,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,148.05) rectangle (121.09,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,148.05) rectangle (124.49,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,148.05) rectangle (127.89,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,148.05) rectangle (131.29,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,148.05) rectangle (134.68,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,148.05) rectangle (138.08,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,148.05) rectangle (141.48,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,148.05) rectangle (144.88,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,148.05) rectangle (148.28,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,148.05) rectangle (151.67,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,148.05) rectangle (155.07,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,148.05) rectangle (158.47,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,148.05) rectangle (161.87,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,148.05) rectangle (165.27,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,148.05) rectangle (168.66,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,148.05) rectangle (172.06,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,148.05) rectangle (175.46,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,148.05) rectangle (178.86,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,148.05) rectangle (182.26,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,148.05) rectangle (185.65,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,148.05) rectangle (189.05,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,148.05) rectangle (192.45,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,148.05) rectangle (195.85,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,148.05) rectangle (199.25,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,148.05) rectangle (202.64,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,148.05) rectangle (206.04,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,148.05) rectangle (209.44,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,148.05) rectangle (212.84,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,148.05) rectangle (216.24,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,148.05) rectangle (219.64,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,148.05) rectangle (223.03,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,148.05) rectangle (226.43,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,148.05) rectangle (229.83,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,148.05) rectangle (233.23,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,148.05) rectangle (236.63,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,148.05) rectangle (240.02,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,148.05) rectangle (243.42,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,148.05) rectangle (246.82,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,148.05) rectangle (250.22,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,148.05) rectangle (253.62,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,148.05) rectangle (257.01,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,148.05) rectangle (260.41,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,148.05) rectangle (263.81,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,148.05) rectangle (267.21,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,148.05) rectangle (270.61,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,148.05) rectangle (274.00,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,148.05) rectangle (277.40,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,148.05) rectangle (280.80,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,148.05) rectangle (284.20,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,148.05) rectangle (287.60,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,148.05) rectangle (290.99,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,148.05) rectangle (294.39,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,148.05) rectangle (297.79,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,148.05) rectangle (301.19,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,148.05) rectangle (304.59,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,148.05) rectangle (307.98,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,148.05) rectangle (311.38,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,148.05) rectangle (314.78,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,148.05) rectangle (318.18,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,148.05) rectangle (321.58,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,148.05) rectangle (324.97,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,148.05) rectangle (328.37,151.45);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,148.05) rectangle (331.77,151.45);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,151.45) rectangle ( 32.74,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,151.45) rectangle ( 36.14,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,151.45) rectangle ( 39.54,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,151.45) rectangle ( 42.94,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,151.45) rectangle ( 46.33,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,151.45) rectangle ( 49.73,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,151.45) rectangle ( 53.13,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,151.45) rectangle ( 56.53,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,151.45) rectangle ( 59.93,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,151.45) rectangle ( 63.32,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,151.45) rectangle ( 66.72,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,151.45) rectangle ( 70.12,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,151.45) rectangle ( 73.52,154.84);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,151.45) rectangle ( 76.92,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,151.45) rectangle ( 80.32,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,151.45) rectangle ( 83.71,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,151.45) rectangle ( 87.11,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,151.45) rectangle ( 90.51,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,151.45) rectangle ( 93.91,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,151.45) rectangle ( 97.31,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,151.45) rectangle (100.70,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,151.45) rectangle (104.10,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,151.45) rectangle (107.50,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,151.45) rectangle (110.90,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,151.45) rectangle (114.30,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,151.45) rectangle (117.69,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,151.45) rectangle (121.09,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,151.45) rectangle (124.49,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,151.45) rectangle (127.89,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,151.45) rectangle (131.29,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,151.45) rectangle (134.68,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,151.45) rectangle (138.08,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,151.45) rectangle (141.48,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,151.45) rectangle (144.88,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,151.45) rectangle (148.28,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,151.45) rectangle (151.67,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,151.45) rectangle (155.07,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,151.45) rectangle (158.47,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,151.45) rectangle (161.87,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,151.45) rectangle (165.27,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,151.45) rectangle (168.66,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,151.45) rectangle (172.06,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,151.45) rectangle (175.46,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,151.45) rectangle (178.86,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,151.45) rectangle (182.26,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,151.45) rectangle (185.65,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,151.45) rectangle (189.05,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,151.45) rectangle (192.45,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,151.45) rectangle (195.85,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,151.45) rectangle (199.25,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,151.45) rectangle (202.64,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,151.45) rectangle (206.04,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,151.45) rectangle (209.44,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,151.45) rectangle (212.84,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,151.45) rectangle (216.24,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,151.45) rectangle (219.64,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,151.45) rectangle (223.03,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,151.45) rectangle (226.43,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,151.45) rectangle (229.83,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,151.45) rectangle (233.23,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,151.45) rectangle (236.63,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,151.45) rectangle (240.02,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,151.45) rectangle (243.42,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,151.45) rectangle (246.82,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,151.45) rectangle (250.22,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,151.45) rectangle (253.62,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,151.45) rectangle (257.01,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,151.45) rectangle (260.41,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,151.45) rectangle (263.81,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,151.45) rectangle (267.21,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,151.45) rectangle (270.61,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,151.45) rectangle (274.00,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,151.45) rectangle (277.40,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,151.45) rectangle (280.80,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,151.45) rectangle (284.20,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,151.45) rectangle (287.60,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,151.45) rectangle (290.99,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,151.45) rectangle (294.39,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,151.45) rectangle (297.79,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,151.45) rectangle (301.19,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,151.45) rectangle (304.59,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,151.45) rectangle (307.98,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,151.45) rectangle (311.38,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,151.45) rectangle (314.78,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,151.45) rectangle (318.18,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,151.45) rectangle (321.58,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,151.45) rectangle (324.97,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,151.45) rectangle (328.37,154.84);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,151.45) rectangle (331.77,154.84);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,154.84) rectangle ( 32.74,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,154.84) rectangle ( 36.14,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,154.84) rectangle ( 39.54,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,154.84) rectangle ( 42.94,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,154.84) rectangle ( 46.33,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,154.84) rectangle ( 49.73,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,154.84) rectangle ( 53.13,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,154.84) rectangle ( 56.53,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,154.84) rectangle ( 59.93,158.24);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,154.84) rectangle ( 63.32,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,154.84) rectangle ( 66.72,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,154.84) rectangle ( 70.12,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,154.84) rectangle ( 73.52,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,154.84) rectangle ( 76.92,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,154.84) rectangle ( 80.32,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,154.84) rectangle ( 83.71,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,154.84) rectangle ( 87.11,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,154.84) rectangle ( 90.51,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,154.84) rectangle ( 93.91,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,154.84) rectangle ( 97.31,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,154.84) rectangle (100.70,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,154.84) rectangle (104.10,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,154.84) rectangle (107.50,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,154.84) rectangle (110.90,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,154.84) rectangle (114.30,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,154.84) rectangle (117.69,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,154.84) rectangle (121.09,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,154.84) rectangle (124.49,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,154.84) rectangle (127.89,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,154.84) rectangle (131.29,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,154.84) rectangle (134.68,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,154.84) rectangle (138.08,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,154.84) rectangle (141.48,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,154.84) rectangle (144.88,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,154.84) rectangle (148.28,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,154.84) rectangle (151.67,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,154.84) rectangle (155.07,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,154.84) rectangle (158.47,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,154.84) rectangle (161.87,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,154.84) rectangle (165.27,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,154.84) rectangle (168.66,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,154.84) rectangle (172.06,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,154.84) rectangle (175.46,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,154.84) rectangle (178.86,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,154.84) rectangle (182.26,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,154.84) rectangle (185.65,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,154.84) rectangle (189.05,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,154.84) rectangle (192.45,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,154.84) rectangle (195.85,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,154.84) rectangle (199.25,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,154.84) rectangle (202.64,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,154.84) rectangle (206.04,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,154.84) rectangle (209.44,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,154.84) rectangle (212.84,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,154.84) rectangle (216.24,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,154.84) rectangle (219.64,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,154.84) rectangle (223.03,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,154.84) rectangle (226.43,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,154.84) rectangle (229.83,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,154.84) rectangle (233.23,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,154.84) rectangle (236.63,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,154.84) rectangle (240.02,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,154.84) rectangle (243.42,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,154.84) rectangle (246.82,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,154.84) rectangle (250.22,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,154.84) rectangle (253.62,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,154.84) rectangle (257.01,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,154.84) rectangle (260.41,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,154.84) rectangle (263.81,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,154.84) rectangle (267.21,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,154.84) rectangle (270.61,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,154.84) rectangle (274.00,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,154.84) rectangle (277.40,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,154.84) rectangle (280.80,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,154.84) rectangle (284.20,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,154.84) rectangle (287.60,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,154.84) rectangle (290.99,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,154.84) rectangle (294.39,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,154.84) rectangle (297.79,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,154.84) rectangle (301.19,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,154.84) rectangle (304.59,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,154.84) rectangle (307.98,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,154.84) rectangle (311.38,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,154.84) rectangle (314.78,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,154.84) rectangle (318.18,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,154.84) rectangle (321.58,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,154.84) rectangle (324.97,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,154.84) rectangle (328.37,158.24);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,154.84) rectangle (331.77,158.24);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,158.24) rectangle ( 32.74,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,158.24) rectangle ( 36.14,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,158.24) rectangle ( 39.54,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,158.24) rectangle ( 42.94,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,158.24) rectangle ( 46.33,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,158.24) rectangle ( 49.73,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,158.24) rectangle ( 53.13,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,158.24) rectangle ( 56.53,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,158.24) rectangle ( 59.93,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,158.24) rectangle ( 63.32,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,158.24) rectangle ( 66.72,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,158.24) rectangle ( 70.12,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,158.24) rectangle ( 73.52,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,158.24) rectangle ( 76.92,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,158.24) rectangle ( 80.32,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,158.24) rectangle ( 83.71,161.64);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,158.24) rectangle ( 87.11,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,158.24) rectangle ( 90.51,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,158.24) rectangle ( 93.91,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,158.24) rectangle ( 97.31,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,158.24) rectangle (100.70,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,158.24) rectangle (104.10,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,158.24) rectangle (107.50,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,158.24) rectangle (110.90,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,158.24) rectangle (114.30,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,158.24) rectangle (117.69,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,158.24) rectangle (121.09,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,158.24) rectangle (124.49,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,158.24) rectangle (127.89,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,158.24) rectangle (131.29,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,158.24) rectangle (134.68,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,158.24) rectangle (138.08,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,158.24) rectangle (141.48,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,158.24) rectangle (144.88,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,158.24) rectangle (148.28,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,158.24) rectangle (151.67,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,158.24) rectangle (155.07,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,158.24) rectangle (158.47,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,158.24) rectangle (161.87,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,158.24) rectangle (165.27,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,158.24) rectangle (168.66,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,158.24) rectangle (172.06,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,158.24) rectangle (175.46,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,158.24) rectangle (178.86,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,158.24) rectangle (182.26,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,158.24) rectangle (185.65,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,158.24) rectangle (189.05,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,158.24) rectangle (192.45,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,158.24) rectangle (195.85,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,158.24) rectangle (199.25,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,158.24) rectangle (202.64,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,158.24) rectangle (206.04,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,158.24) rectangle (209.44,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,158.24) rectangle (212.84,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,158.24) rectangle (216.24,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,158.24) rectangle (219.64,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,158.24) rectangle (223.03,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,158.24) rectangle (226.43,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,158.24) rectangle (229.83,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,158.24) rectangle (233.23,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,158.24) rectangle (236.63,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,158.24) rectangle (240.02,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,158.24) rectangle (243.42,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,158.24) rectangle (246.82,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,158.24) rectangle (250.22,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,158.24) rectangle (253.62,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,158.24) rectangle (257.01,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,158.24) rectangle (260.41,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,158.24) rectangle (263.81,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,158.24) rectangle (267.21,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,158.24) rectangle (270.61,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,158.24) rectangle (274.00,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,158.24) rectangle (277.40,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,158.24) rectangle (280.80,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,158.24) rectangle (284.20,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,158.24) rectangle (287.60,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,158.24) rectangle (290.99,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,158.24) rectangle (294.39,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,158.24) rectangle (297.79,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,158.24) rectangle (301.19,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,158.24) rectangle (304.59,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,158.24) rectangle (307.98,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,158.24) rectangle (311.38,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,158.24) rectangle (314.78,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,158.24) rectangle (318.18,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,158.24) rectangle (321.58,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,158.24) rectangle (324.97,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,158.24) rectangle (328.37,161.64);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,158.24) rectangle (331.77,161.64);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,161.64) rectangle ( 32.74,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,161.64) rectangle ( 36.14,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,161.64) rectangle ( 39.54,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,161.64) rectangle ( 42.94,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,161.64) rectangle ( 46.33,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,161.64) rectangle ( 49.73,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,161.64) rectangle ( 53.13,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,161.64) rectangle ( 56.53,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,161.64) rectangle ( 59.93,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,161.64) rectangle ( 63.32,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,161.64) rectangle ( 66.72,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,161.64) rectangle ( 70.12,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,161.64) rectangle ( 73.52,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,161.64) rectangle ( 76.92,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,161.64) rectangle ( 80.32,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,161.64) rectangle ( 83.71,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,161.64) rectangle ( 87.11,165.04);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,161.64) rectangle ( 90.51,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,161.64) rectangle ( 93.91,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,161.64) rectangle ( 97.31,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,161.64) rectangle (100.70,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,161.64) rectangle (104.10,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,161.64) rectangle (107.50,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,161.64) rectangle (110.90,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,161.64) rectangle (114.30,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,161.64) rectangle (117.69,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,161.64) rectangle (121.09,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,161.64) rectangle (124.49,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,161.64) rectangle (127.89,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,161.64) rectangle (131.29,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,161.64) rectangle (134.68,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,161.64) rectangle (138.08,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,161.64) rectangle (141.48,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,161.64) rectangle (144.88,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,161.64) rectangle (148.28,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,161.64) rectangle (151.67,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,161.64) rectangle (155.07,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,161.64) rectangle (158.47,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,161.64) rectangle (161.87,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,161.64) rectangle (165.27,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,161.64) rectangle (168.66,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,161.64) rectangle (172.06,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,161.64) rectangle (175.46,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,161.64) rectangle (178.86,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,161.64) rectangle (182.26,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,161.64) rectangle (185.65,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,161.64) rectangle (189.05,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,161.64) rectangle (192.45,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,161.64) rectangle (195.85,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,161.64) rectangle (199.25,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,161.64) rectangle (202.64,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,161.64) rectangle (206.04,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,161.64) rectangle (209.44,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,161.64) rectangle (212.84,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,161.64) rectangle (216.24,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,161.64) rectangle (219.64,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,161.64) rectangle (223.03,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,161.64) rectangle (226.43,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,161.64) rectangle (229.83,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,161.64) rectangle (233.23,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,161.64) rectangle (236.63,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,161.64) rectangle (240.02,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,161.64) rectangle (243.42,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,161.64) rectangle (246.82,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,161.64) rectangle (250.22,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,161.64) rectangle (253.62,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,161.64) rectangle (257.01,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,161.64) rectangle (260.41,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,161.64) rectangle (263.81,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,161.64) rectangle (267.21,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,161.64) rectangle (270.61,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,161.64) rectangle (274.00,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,161.64) rectangle (277.40,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,161.64) rectangle (280.80,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,161.64) rectangle (284.20,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,161.64) rectangle (287.60,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,161.64) rectangle (290.99,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,161.64) rectangle (294.39,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,161.64) rectangle (297.79,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,161.64) rectangle (301.19,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,161.64) rectangle (304.59,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,161.64) rectangle (307.98,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,161.64) rectangle (311.38,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,161.64) rectangle (314.78,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,161.64) rectangle (318.18,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,161.64) rectangle (321.58,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,161.64) rectangle (324.97,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,161.64) rectangle (328.37,165.04);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,161.64) rectangle (331.77,165.04);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,165.04) rectangle ( 32.74,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,165.04) rectangle ( 36.14,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,165.04) rectangle ( 39.54,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,165.04) rectangle ( 42.94,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,165.04) rectangle ( 46.33,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,165.04) rectangle ( 49.73,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,165.04) rectangle ( 53.13,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,165.04) rectangle ( 56.53,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,165.04) rectangle ( 59.93,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,165.04) rectangle ( 63.32,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,165.04) rectangle ( 66.72,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,165.04) rectangle ( 70.12,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,165.04) rectangle ( 73.52,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,165.04) rectangle ( 76.92,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,165.04) rectangle ( 80.32,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,165.04) rectangle ( 83.71,168.44);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,165.04) rectangle ( 87.11,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,165.04) rectangle ( 90.51,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,165.04) rectangle ( 93.91,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,165.04) rectangle ( 97.31,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,165.04) rectangle (100.70,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,165.04) rectangle (104.10,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,165.04) rectangle (107.50,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,165.04) rectangle (110.90,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,165.04) rectangle (114.30,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,165.04) rectangle (117.69,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,165.04) rectangle (121.09,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,165.04) rectangle (124.49,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,165.04) rectangle (127.89,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,165.04) rectangle (131.29,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,165.04) rectangle (134.68,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,165.04) rectangle (138.08,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,165.04) rectangle (141.48,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,165.04) rectangle (144.88,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,165.04) rectangle (148.28,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,165.04) rectangle (151.67,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,165.04) rectangle (155.07,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,165.04) rectangle (158.47,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,165.04) rectangle (161.87,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,165.04) rectangle (165.27,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,165.04) rectangle (168.66,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,165.04) rectangle (172.06,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,165.04) rectangle (175.46,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,165.04) rectangle (178.86,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,165.04) rectangle (182.26,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,165.04) rectangle (185.65,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,165.04) rectangle (189.05,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,165.04) rectangle (192.45,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,165.04) rectangle (195.85,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,165.04) rectangle (199.25,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,165.04) rectangle (202.64,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,165.04) rectangle (206.04,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,165.04) rectangle (209.44,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,165.04) rectangle (212.84,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,165.04) rectangle (216.24,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,165.04) rectangle (219.64,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,165.04) rectangle (223.03,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,165.04) rectangle (226.43,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,165.04) rectangle (229.83,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,165.04) rectangle (233.23,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,165.04) rectangle (236.63,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,165.04) rectangle (240.02,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,165.04) rectangle (243.42,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,165.04) rectangle (246.82,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,165.04) rectangle (250.22,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,165.04) rectangle (253.62,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,165.04) rectangle (257.01,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,165.04) rectangle (260.41,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,165.04) rectangle (263.81,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,165.04) rectangle (267.21,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,165.04) rectangle (270.61,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,165.04) rectangle (274.00,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,165.04) rectangle (277.40,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,165.04) rectangle (280.80,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,165.04) rectangle (284.20,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,165.04) rectangle (287.60,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,165.04) rectangle (290.99,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,165.04) rectangle (294.39,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,165.04) rectangle (297.79,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,165.04) rectangle (301.19,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,165.04) rectangle (304.59,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,165.04) rectangle (307.98,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,165.04) rectangle (311.38,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,165.04) rectangle (314.78,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,165.04) rectangle (318.18,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,165.04) rectangle (321.58,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,165.04) rectangle (324.97,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,165.04) rectangle (328.37,168.44);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,165.04) rectangle (331.77,168.44);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,168.44) rectangle ( 32.74,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,168.44) rectangle ( 36.14,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,168.44) rectangle ( 39.54,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,168.44) rectangle ( 42.94,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,168.44) rectangle ( 46.33,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,168.44) rectangle ( 49.73,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,168.44) rectangle ( 53.13,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,168.44) rectangle ( 56.53,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,168.44) rectangle ( 59.93,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,168.44) rectangle ( 63.32,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,168.44) rectangle ( 66.72,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,168.44) rectangle ( 70.12,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,168.44) rectangle ( 73.52,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,168.44) rectangle ( 76.92,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,168.44) rectangle ( 80.32,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,168.44) rectangle ( 83.71,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,168.44) rectangle ( 87.11,171.83);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,168.44) rectangle ( 90.51,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,168.44) rectangle ( 93.91,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,168.44) rectangle ( 97.31,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,168.44) rectangle (100.70,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,168.44) rectangle (104.10,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,168.44) rectangle (107.50,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,168.44) rectangle (110.90,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,168.44) rectangle (114.30,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,168.44) rectangle (117.69,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,168.44) rectangle (121.09,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,168.44) rectangle (124.49,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,168.44) rectangle (127.89,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,168.44) rectangle (131.29,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,168.44) rectangle (134.68,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,168.44) rectangle (138.08,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,168.44) rectangle (141.48,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,168.44) rectangle (144.88,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,168.44) rectangle (148.28,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,168.44) rectangle (151.67,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,168.44) rectangle (155.07,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,168.44) rectangle (158.47,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,168.44) rectangle (161.87,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,168.44) rectangle (165.27,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,168.44) rectangle (168.66,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,168.44) rectangle (172.06,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,168.44) rectangle (175.46,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,168.44) rectangle (178.86,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,168.44) rectangle (182.26,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,168.44) rectangle (185.65,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,168.44) rectangle (189.05,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,168.44) rectangle (192.45,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,168.44) rectangle (195.85,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,168.44) rectangle (199.25,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,168.44) rectangle (202.64,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,168.44) rectangle (206.04,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,168.44) rectangle (209.44,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,168.44) rectangle (212.84,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,168.44) rectangle (216.24,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,168.44) rectangle (219.64,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,168.44) rectangle (223.03,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,168.44) rectangle (226.43,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,168.44) rectangle (229.83,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,168.44) rectangle (233.23,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,168.44) rectangle (236.63,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,168.44) rectangle (240.02,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,168.44) rectangle (243.42,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,168.44) rectangle (246.82,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,168.44) rectangle (250.22,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,168.44) rectangle (253.62,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,168.44) rectangle (257.01,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,168.44) rectangle (260.41,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,168.44) rectangle (263.81,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,168.44) rectangle (267.21,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,168.44) rectangle (270.61,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,168.44) rectangle (274.00,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,168.44) rectangle (277.40,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,168.44) rectangle (280.80,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,168.44) rectangle (284.20,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,168.44) rectangle (287.60,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,168.44) rectangle (290.99,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,168.44) rectangle (294.39,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,168.44) rectangle (297.79,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,168.44) rectangle (301.19,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,168.44) rectangle (304.59,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,168.44) rectangle (307.98,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,168.44) rectangle (311.38,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,168.44) rectangle (314.78,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,168.44) rectangle (318.18,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,168.44) rectangle (321.58,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,168.44) rectangle (324.97,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,168.44) rectangle (328.37,171.83);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,168.44) rectangle (331.77,171.83);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,171.83) rectangle ( 32.74,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,171.83) rectangle ( 36.14,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,171.83) rectangle ( 39.54,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,171.83) rectangle ( 42.94,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,171.83) rectangle ( 46.33,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,171.83) rectangle ( 49.73,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,171.83) rectangle ( 53.13,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,171.83) rectangle ( 56.53,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,171.83) rectangle ( 59.93,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,171.83) rectangle ( 63.32,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,171.83) rectangle ( 66.72,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,171.83) rectangle ( 70.12,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,171.83) rectangle ( 73.52,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,171.83) rectangle ( 76.92,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,171.83) rectangle ( 80.32,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,171.83) rectangle ( 83.71,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,171.83) rectangle ( 87.11,175.23);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,171.83) rectangle ( 90.51,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,171.83) rectangle ( 93.91,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,171.83) rectangle ( 97.31,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,171.83) rectangle (100.70,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,171.83) rectangle (104.10,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,171.83) rectangle (107.50,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,171.83) rectangle (110.90,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,171.83) rectangle (114.30,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,171.83) rectangle (117.69,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,171.83) rectangle (121.09,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,171.83) rectangle (124.49,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,171.83) rectangle (127.89,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,171.83) rectangle (131.29,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,171.83) rectangle (134.68,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,171.83) rectangle (138.08,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,171.83) rectangle (141.48,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,171.83) rectangle (144.88,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,171.83) rectangle (148.28,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,171.83) rectangle (151.67,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,171.83) rectangle (155.07,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,171.83) rectangle (158.47,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,171.83) rectangle (161.87,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,171.83) rectangle (165.27,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,171.83) rectangle (168.66,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,171.83) rectangle (172.06,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,171.83) rectangle (175.46,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,171.83) rectangle (178.86,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,171.83) rectangle (182.26,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,171.83) rectangle (185.65,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,171.83) rectangle (189.05,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,171.83) rectangle (192.45,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,171.83) rectangle (195.85,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,171.83) rectangle (199.25,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,171.83) rectangle (202.64,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,171.83) rectangle (206.04,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,171.83) rectangle (209.44,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,171.83) rectangle (212.84,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,171.83) rectangle (216.24,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,171.83) rectangle (219.64,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,171.83) rectangle (223.03,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,171.83) rectangle (226.43,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,171.83) rectangle (229.83,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,171.83) rectangle (233.23,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,171.83) rectangle (236.63,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,171.83) rectangle (240.02,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,171.83) rectangle (243.42,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,171.83) rectangle (246.82,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,171.83) rectangle (250.22,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,171.83) rectangle (253.62,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,171.83) rectangle (257.01,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,171.83) rectangle (260.41,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,171.83) rectangle (263.81,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,171.83) rectangle (267.21,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,171.83) rectangle (270.61,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,171.83) rectangle (274.00,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,171.83) rectangle (277.40,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,171.83) rectangle (280.80,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,171.83) rectangle (284.20,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,171.83) rectangle (287.60,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,171.83) rectangle (290.99,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,171.83) rectangle (294.39,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,171.83) rectangle (297.79,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,171.83) rectangle (301.19,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,171.83) rectangle (304.59,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,171.83) rectangle (307.98,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,171.83) rectangle (311.38,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,171.83) rectangle (314.78,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,171.83) rectangle (318.18,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,171.83) rectangle (321.58,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,171.83) rectangle (324.97,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,171.83) rectangle (328.37,175.23);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,171.83) rectangle (331.77,175.23);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,175.23) rectangle ( 32.74,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,175.23) rectangle ( 36.14,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,175.23) rectangle ( 39.54,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,175.23) rectangle ( 42.94,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,175.23) rectangle ( 46.33,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,175.23) rectangle ( 49.73,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,175.23) rectangle ( 53.13,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,175.23) rectangle ( 56.53,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,175.23) rectangle ( 59.93,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,175.23) rectangle ( 63.32,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,175.23) rectangle ( 66.72,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,175.23) rectangle ( 70.12,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,175.23) rectangle ( 73.52,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,175.23) rectangle ( 76.92,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,175.23) rectangle ( 80.32,178.63);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,175.23) rectangle ( 83.71,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,175.23) rectangle ( 87.11,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,175.23) rectangle ( 90.51,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,175.23) rectangle ( 93.91,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,175.23) rectangle ( 97.31,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,175.23) rectangle (100.70,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,175.23) rectangle (104.10,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,175.23) rectangle (107.50,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,175.23) rectangle (110.90,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,175.23) rectangle (114.30,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,175.23) rectangle (117.69,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,175.23) rectangle (121.09,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,175.23) rectangle (124.49,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,175.23) rectangle (127.89,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,175.23) rectangle (131.29,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,175.23) rectangle (134.68,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,175.23) rectangle (138.08,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,175.23) rectangle (141.48,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,175.23) rectangle (144.88,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,175.23) rectangle (148.28,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,175.23) rectangle (151.67,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,175.23) rectangle (155.07,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,175.23) rectangle (158.47,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,175.23) rectangle (161.87,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,175.23) rectangle (165.27,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,175.23) rectangle (168.66,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,175.23) rectangle (172.06,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,175.23) rectangle (175.46,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,175.23) rectangle (178.86,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,175.23) rectangle (182.26,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,175.23) rectangle (185.65,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,175.23) rectangle (189.05,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,175.23) rectangle (192.45,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,175.23) rectangle (195.85,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,175.23) rectangle (199.25,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,175.23) rectangle (202.64,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,175.23) rectangle (206.04,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,175.23) rectangle (209.44,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,175.23) rectangle (212.84,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,175.23) rectangle (216.24,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,175.23) rectangle (219.64,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,175.23) rectangle (223.03,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,175.23) rectangle (226.43,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,175.23) rectangle (229.83,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,175.23) rectangle (233.23,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,175.23) rectangle (236.63,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,175.23) rectangle (240.02,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,175.23) rectangle (243.42,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,175.23) rectangle (246.82,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,175.23) rectangle (250.22,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,175.23) rectangle (253.62,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,175.23) rectangle (257.01,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,175.23) rectangle (260.41,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,175.23) rectangle (263.81,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,175.23) rectangle (267.21,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,175.23) rectangle (270.61,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,175.23) rectangle (274.00,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,175.23) rectangle (277.40,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,175.23) rectangle (280.80,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,175.23) rectangle (284.20,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,175.23) rectangle (287.60,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,175.23) rectangle (290.99,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,175.23) rectangle (294.39,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,175.23) rectangle (297.79,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,175.23) rectangle (301.19,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,175.23) rectangle (304.59,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,175.23) rectangle (307.98,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,175.23) rectangle (311.38,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,175.23) rectangle (314.78,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,175.23) rectangle (318.18,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,175.23) rectangle (321.58,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,175.23) rectangle (324.97,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,175.23) rectangle (328.37,178.63);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,175.23) rectangle (331.77,178.63);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,178.63) rectangle ( 32.74,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,178.63) rectangle ( 36.14,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,178.63) rectangle ( 39.54,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,178.63) rectangle ( 42.94,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,178.63) rectangle ( 46.33,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,178.63) rectangle ( 49.73,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,178.63) rectangle ( 53.13,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,178.63) rectangle ( 56.53,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,178.63) rectangle ( 59.93,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,178.63) rectangle ( 63.32,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,178.63) rectangle ( 66.72,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,178.63) rectangle ( 70.12,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,178.63) rectangle ( 73.52,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,178.63) rectangle ( 76.92,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,178.63) rectangle ( 80.32,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,178.63) rectangle ( 83.71,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,178.63) rectangle ( 87.11,182.03);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,178.63) rectangle ( 90.51,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,178.63) rectangle ( 93.91,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,178.63) rectangle ( 97.31,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,178.63) rectangle (100.70,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,178.63) rectangle (104.10,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,178.63) rectangle (107.50,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,178.63) rectangle (110.90,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,178.63) rectangle (114.30,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,178.63) rectangle (117.69,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,178.63) rectangle (121.09,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,178.63) rectangle (124.49,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,178.63) rectangle (127.89,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,178.63) rectangle (131.29,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,178.63) rectangle (134.68,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,178.63) rectangle (138.08,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,178.63) rectangle (141.48,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,178.63) rectangle (144.88,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,178.63) rectangle (148.28,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,178.63) rectangle (151.67,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,178.63) rectangle (155.07,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,178.63) rectangle (158.47,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,178.63) rectangle (161.87,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,178.63) rectangle (165.27,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,178.63) rectangle (168.66,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,178.63) rectangle (172.06,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,178.63) rectangle (175.46,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,178.63) rectangle (178.86,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,178.63) rectangle (182.26,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,178.63) rectangle (185.65,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,178.63) rectangle (189.05,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,178.63) rectangle (192.45,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,178.63) rectangle (195.85,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,178.63) rectangle (199.25,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,178.63) rectangle (202.64,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,178.63) rectangle (206.04,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,178.63) rectangle (209.44,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,178.63) rectangle (212.84,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,178.63) rectangle (216.24,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,178.63) rectangle (219.64,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,178.63) rectangle (223.03,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,178.63) rectangle (226.43,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,178.63) rectangle (229.83,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,178.63) rectangle (233.23,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,178.63) rectangle (236.63,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,178.63) rectangle (240.02,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,178.63) rectangle (243.42,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,178.63) rectangle (246.82,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,178.63) rectangle (250.22,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,178.63) rectangle (253.62,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,178.63) rectangle (257.01,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,178.63) rectangle (260.41,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,178.63) rectangle (263.81,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,178.63) rectangle (267.21,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,178.63) rectangle (270.61,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,178.63) rectangle (274.00,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,178.63) rectangle (277.40,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,178.63) rectangle (280.80,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,178.63) rectangle (284.20,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,178.63) rectangle (287.60,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,178.63) rectangle (290.99,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,178.63) rectangle (294.39,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,178.63) rectangle (297.79,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,178.63) rectangle (301.19,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,178.63) rectangle (304.59,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,178.63) rectangle (307.98,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,178.63) rectangle (311.38,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,178.63) rectangle (314.78,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,178.63) rectangle (318.18,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,178.63) rectangle (321.58,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,178.63) rectangle (324.97,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,178.63) rectangle (328.37,182.03);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,178.63) rectangle (331.77,182.03);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,182.03) rectangle ( 32.74,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,182.03) rectangle ( 36.14,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,182.03) rectangle ( 39.54,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,182.03) rectangle ( 42.94,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,182.03) rectangle ( 46.33,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,182.03) rectangle ( 49.73,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,182.03) rectangle ( 53.13,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,182.03) rectangle ( 56.53,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,182.03) rectangle ( 59.93,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,182.03) rectangle ( 63.32,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,182.03) rectangle ( 66.72,185.43);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,182.03) rectangle ( 70.12,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,182.03) rectangle ( 73.52,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,182.03) rectangle ( 76.92,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,182.03) rectangle ( 80.32,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,182.03) rectangle ( 83.71,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,182.03) rectangle ( 87.11,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,182.03) rectangle ( 90.51,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,182.03) rectangle ( 93.91,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,182.03) rectangle ( 97.31,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,182.03) rectangle (100.70,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,182.03) rectangle (104.10,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,182.03) rectangle (107.50,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,182.03) rectangle (110.90,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,182.03) rectangle (114.30,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,182.03) rectangle (117.69,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,182.03) rectangle (121.09,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,182.03) rectangle (124.49,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,182.03) rectangle (127.89,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,182.03) rectangle (131.29,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,182.03) rectangle (134.68,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,182.03) rectangle (138.08,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,182.03) rectangle (141.48,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,182.03) rectangle (144.88,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,182.03) rectangle (148.28,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,182.03) rectangle (151.67,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,182.03) rectangle (155.07,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,182.03) rectangle (158.47,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,182.03) rectangle (161.87,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,182.03) rectangle (165.27,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,182.03) rectangle (168.66,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,182.03) rectangle (172.06,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,182.03) rectangle (175.46,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,182.03) rectangle (178.86,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,182.03) rectangle (182.26,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,182.03) rectangle (185.65,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,182.03) rectangle (189.05,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,182.03) rectangle (192.45,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,182.03) rectangle (195.85,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,182.03) rectangle (199.25,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,182.03) rectangle (202.64,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,182.03) rectangle (206.04,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,182.03) rectangle (209.44,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,182.03) rectangle (212.84,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,182.03) rectangle (216.24,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,182.03) rectangle (219.64,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,182.03) rectangle (223.03,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,182.03) rectangle (226.43,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,182.03) rectangle (229.83,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,182.03) rectangle (233.23,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,182.03) rectangle (236.63,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,182.03) rectangle (240.02,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,182.03) rectangle (243.42,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,182.03) rectangle (246.82,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,182.03) rectangle (250.22,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,182.03) rectangle (253.62,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,182.03) rectangle (257.01,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,182.03) rectangle (260.41,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,182.03) rectangle (263.81,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,182.03) rectangle (267.21,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,182.03) rectangle (270.61,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,182.03) rectangle (274.00,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,182.03) rectangle (277.40,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,182.03) rectangle (280.80,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,182.03) rectangle (284.20,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,182.03) rectangle (287.60,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,182.03) rectangle (290.99,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,182.03) rectangle (294.39,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,182.03) rectangle (297.79,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,182.03) rectangle (301.19,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,182.03) rectangle (304.59,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,182.03) rectangle (307.98,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,182.03) rectangle (311.38,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,182.03) rectangle (314.78,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,182.03) rectangle (318.18,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,182.03) rectangle (321.58,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,182.03) rectangle (324.97,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,182.03) rectangle (328.37,185.43);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,182.03) rectangle (331.77,185.43);
\definecolor{fillColor}{RGB}{54,100,139}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 29.34,185.43) rectangle ( 32.74,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 32.74,185.43) rectangle ( 36.14,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 36.14,185.43) rectangle ( 39.54,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 39.54,185.43) rectangle ( 42.94,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 42.94,185.43) rectangle ( 46.33,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 46.33,185.43) rectangle ( 49.73,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 49.73,185.43) rectangle ( 53.13,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 53.13,185.43) rectangle ( 56.53,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 56.53,185.43) rectangle ( 59.93,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 59.93,185.43) rectangle ( 63.32,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 63.32,185.43) rectangle ( 66.72,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 66.72,185.43) rectangle ( 70.12,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 70.12,185.43) rectangle ( 73.52,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 73.52,185.43) rectangle ( 76.92,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 76.92,185.43) rectangle ( 80.32,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 80.32,185.43) rectangle ( 83.71,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 83.71,185.43) rectangle ( 87.11,188.82);
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 87.11,185.43) rectangle ( 90.51,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 90.51,185.43) rectangle ( 93.91,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 93.91,185.43) rectangle ( 97.31,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] ( 97.31,185.43) rectangle (100.70,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (100.70,185.43) rectangle (104.10,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (104.10,185.43) rectangle (107.50,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (107.50,185.43) rectangle (110.90,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (110.90,185.43) rectangle (114.30,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (114.30,185.43) rectangle (117.69,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (117.69,185.43) rectangle (121.09,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (121.09,185.43) rectangle (124.49,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (124.49,185.43) rectangle (127.89,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (127.89,185.43) rectangle (131.29,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (131.29,185.43) rectangle (134.68,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (134.68,185.43) rectangle (138.08,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (138.08,185.43) rectangle (141.48,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (141.48,185.43) rectangle (144.88,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (144.88,185.43) rectangle (148.28,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (148.28,185.43) rectangle (151.67,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (151.67,185.43) rectangle (155.07,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (155.07,185.43) rectangle (158.47,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (158.47,185.43) rectangle (161.87,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (161.87,185.43) rectangle (165.27,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (165.27,185.43) rectangle (168.66,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (168.66,185.43) rectangle (172.06,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (172.06,185.43) rectangle (175.46,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (175.46,185.43) rectangle (178.86,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (178.86,185.43) rectangle (182.26,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (182.26,185.43) rectangle (185.65,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (185.65,185.43) rectangle (189.05,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (189.05,185.43) rectangle (192.45,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (192.45,185.43) rectangle (195.85,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (195.85,185.43) rectangle (199.25,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (199.25,185.43) rectangle (202.64,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (202.64,185.43) rectangle (206.04,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (206.04,185.43) rectangle (209.44,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (209.44,185.43) rectangle (212.84,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (212.84,185.43) rectangle (216.24,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (216.24,185.43) rectangle (219.64,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (219.64,185.43) rectangle (223.03,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (223.03,185.43) rectangle (226.43,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (226.43,185.43) rectangle (229.83,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (229.83,185.43) rectangle (233.23,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (233.23,185.43) rectangle (236.63,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (236.63,185.43) rectangle (240.02,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (240.02,185.43) rectangle (243.42,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (243.42,185.43) rectangle (246.82,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (246.82,185.43) rectangle (250.22,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (250.22,185.43) rectangle (253.62,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (253.62,185.43) rectangle (257.01,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (257.01,185.43) rectangle (260.41,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (260.41,185.43) rectangle (263.81,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (263.81,185.43) rectangle (267.21,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (267.21,185.43) rectangle (270.61,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (270.61,185.43) rectangle (274.00,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (274.00,185.43) rectangle (277.40,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (277.40,185.43) rectangle (280.80,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (280.80,185.43) rectangle (284.20,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (284.20,185.43) rectangle (287.60,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (287.60,185.43) rectangle (290.99,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (290.99,185.43) rectangle (294.39,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (294.39,185.43) rectangle (297.79,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (297.79,185.43) rectangle (301.19,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (301.19,185.43) rectangle (304.59,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (304.59,185.43) rectangle (307.98,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (307.98,185.43) rectangle (311.38,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (311.38,185.43) rectangle (314.78,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (314.78,185.43) rectangle (318.18,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (318.18,185.43) rectangle (321.58,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (321.58,185.43) rectangle (324.97,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (324.97,185.43) rectangle (328.37,188.82);

\path[draw=drawColor,line width= 0.1pt,line cap=rect,fill=fillColor] (328.37,185.43) rectangle (331.77,188.82);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (339.67,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 21.90,120.08) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 21.90,133.67) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 21.90,150.66) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 21.90,167.66) {15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 21.90,184.65) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (339.67,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 31.04,108.46) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 61.63,108.46) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 95.61,108.46) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (129.59,108.46) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (163.57,108.46) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (197.55,108.46) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (231.53,108.46) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (265.51,108.46) {70};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (299.49,108.46) {80};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (333.47,108.46) {90};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (339.67,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (180.56, 98.61) {Step};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (339.67,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,154.84) {Predictor};
\end{scope}
\end{tikzpicture}

\end{document}