### 2021-05-12 paper/figures/simulateddata-timings.tex

% Created by tikzDevice version 0.12.3.1 on 2021-05-11 11:28:06
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,144.54);
\begin{scope}
\path[clip] ( 27.95, 25.11) rectangle (245.47,140.04);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 25.15) --
	(245.47, 25.15);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 53.21) --
	(245.47, 53.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 81.27) --
	(245.47, 81.27);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95,109.33) --
	(245.47,109.33);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95,137.39) --
	(245.47,137.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 39.18) --
	(245.47, 39.18);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 67.24) --
	(245.47, 67.24);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 95.30) --
	(245.47, 95.30);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95,123.36) --
	(245.47,123.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 68.73, 25.11) --
	( 68.73,140.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (136.71, 25.11) --
	(136.71,140.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (204.68, 25.11) --
	(204.68,140.04);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 35.59) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 30.33) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 69.35) -- ( 55.99, 81.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 56.03) -- ( 55.99, 36.07);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 44.52, 69.35) --
	( 44.52, 56.03) --
	( 67.46, 56.03) --
	( 67.46, 69.35) --
	( 44.52, 69.35) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 44.52, 64.82) -- ( 67.46, 64.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48,100.56) -- ( 81.48,134.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48, 75.58) -- ( 81.48, 47.67);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 70.01,100.56) --
	( 70.01, 75.58) --
	( 92.95, 75.58) --
	( 92.95,100.56) --
	( 70.01,100.56) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 70.01, 87.94) -- ( 92.95, 87.94);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 40.25) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 48.90) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 46.58) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 43.27) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 73.30) -- (123.96, 85.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 63.82) -- (123.96, 55.13);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (112.49, 73.30) --
	(112.49, 63.82) --
	(135.43, 63.82) --
	(135.43, 73.30) --
	(112.49, 73.30) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (112.49, 68.49) -- (135.43, 68.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45,106.12) -- (149.45,122.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45, 87.88) -- (149.45, 62.14);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (137.98,106.12) --
	(137.98, 87.88) --
	(160.92, 87.88) --
	(160.92,106.12) --
	(137.98,106.12) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.98, 99.28) -- (160.92, 99.28);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (191.94, 89.45) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (191.94, 90.50) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (191.94, 53.69) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 73.09) -- (191.94, 80.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 67.49) -- (191.94, 60.23);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (180.47, 73.09) --
	(180.47, 67.49) --
	(203.41, 67.49) --
	(203.41, 73.09) --
	(180.47, 73.09) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (180.47, 71.06) -- (203.41, 71.06);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (217.43,132.84) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43,114.32) -- (217.43,122.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43,103.13) -- (217.43, 86.49);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (205.96,114.32) --
	(205.96,103.13) --
	(228.90,103.13) --
	(228.90,114.32) --
	(205.96,114.32) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (205.96,109.09) -- (228.90,109.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 36.70) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 64.76) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 92.82) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90,120.88) {3.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 68.73, 16.10) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (136.71, 16.10) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (204.68, 16.10) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (136.71,  6.25) {Signal-to-Noise Ratio};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70, 82.57) {Time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 81.77) --
	(266.20, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 91.16) --
	(266.20, 93.33);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 83.94) rectangle (271.62, 91.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 87.55) --
	(271.62, 87.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 67.31) --
	(266.20, 69.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 76.71) --
	(266.20, 78.88);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 69.48) rectangle (271.62, 76.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 73.10) --
	(271.62, 73.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 85.07) {Look-Ahead};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 70.62) {Standard};
\end{scope}
\end{tikzpicture}

\end{document}