### 2021-05-12 paper/figures/simulateddata-timings.tex
Update results and figures

% Created by tikzDevice version 0.12.3.1 on 2021-05-12 11:19:34
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,144.54);
\begin{scope}
\path[clip] ( 27.95, 25.11) rectangle (245.47,140.04);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 25.98) --
	(245.47, 25.98);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 57.72) --
	(245.47, 57.72);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 89.47) --
	(245.47, 89.47);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95,121.21) --
	(245.47,121.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 41.85) --
	(245.47, 41.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 73.59) --
	(245.47, 73.59);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95,105.34) --
	(245.47,105.34);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95,137.08) --
	(245.47,137.08);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 68.73, 25.11) --
	( 68.73,140.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (136.71, 25.11) --
	(136.71,140.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (204.68, 25.11) --
	(204.68,140.04);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 30.73) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 37.51) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 31.21) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 30.33) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 67.44) -- ( 55.99, 76.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 55.70) -- ( 55.99, 42.49);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 44.52, 67.44) --
	( 44.52, 55.70) --
	( 67.46, 55.70) --
	( 67.46, 67.44) --
	( 44.52, 67.44) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 44.52, 63.76) -- ( 67.46, 63.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48, 96.61) -- ( 81.48,133.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48, 71.28) -- ( 81.48, 52.00);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 70.01, 96.61) --
	( 70.01, 71.28) --
	( 92.95, 71.28) --
	( 92.95, 96.61) --
	( 70.01, 96.61) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 70.01, 82.83) -- ( 92.95, 82.83);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 50.12) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 43.36) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 55.84) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 89.07) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 45.41) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 48.98) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 48.74) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 39.67) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 71.37) -- (123.96, 76.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 65.29) -- (123.96, 60.68);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (112.49, 71.37) --
	(112.49, 65.29) --
	(135.43, 65.29) --
	(135.43, 71.37) --
	(112.49, 71.37) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (112.49, 67.86) -- (135.43, 67.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45,106.97) -- (149.45,134.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45, 88.51) -- (149.45, 73.42);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (137.98,106.97) --
	(137.98, 88.51) --
	(160.92, 88.51) --
	(160.92,106.97) --
	(137.98,106.97) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.98,100.56) -- (160.92,100.56);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (191.94, 94.02) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 76.40) -- (191.94, 86.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 68.57) -- (191.94, 59.83);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (180.47, 76.40) --
	(180.47, 68.57) --
	(203.41, 68.57) --
	(203.41, 76.40) --
	(180.47, 76.40) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (180.47, 72.60) -- (203.41, 72.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43,119.36) -- (217.43,134.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43,102.67) -- (217.43, 79.18);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (205.96,119.36) --
	(205.96,102.67) --
	(228.90,102.67) --
	(228.90,119.36) --
	(205.96,119.36) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (205.96,112.22) -- (228.90,112.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 39.37) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 71.11) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90,102.86) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90,134.60) {3.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 68.73, 16.10) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (136.71, 16.10) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (204.68, 16.10) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (136.71,  6.25) {Signal-to-Noise Ratio};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70, 82.57) {Time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 81.77) --
	(266.20, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 91.16) --
	(266.20, 93.33);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 83.94) rectangle (271.62, 91.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 87.55) --
	(271.62, 87.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 67.31) --
	(266.20, 69.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 76.71) --
	(266.20, 78.88);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 69.48) rectangle (271.62, 76.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 73.10) --
	(271.62, 73.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 85.07) {Look-Ahead};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 70.62) {Standard};
\end{scope}
\end{tikzpicture}

\end{document}