### 2021-06-24 paper/figures/simulated-data-timings.tex
Revise paper based on reviewer comments

% Created by tikzDevice version 0.12.3.1 on 2021-06-24 15:58:06
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,130.09);
\begin{scope}
\path[clip] ( 27.95, 25.11) rectangle (245.47,125.59);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 25.87) --
	(245.47, 25.87);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 53.62) --
	(245.47, 53.62);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95, 81.37) --
	(245.47, 81.37);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 27.95,109.12) --
	(245.47,109.12);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 39.75) --
	(245.47, 39.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 67.50) --
	(245.47, 67.50);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95, 95.25) --
	(245.47, 95.25);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.95,123.00) --
	(245.47,123.00);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 68.73, 25.11) --
	( 68.73,125.59);

\path[draw=drawColor,line width= 0.5pt,line join=round] (136.71, 25.11) --
	(136.71,125.59);

\path[draw=drawColor,line width= 0.5pt,line join=round] (204.68, 25.11) --
	(204.68,125.59);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 30.02) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 35.95) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 30.44) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] ( 55.99, 29.67) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 62.12) -- ( 55.99, 70.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.99, 51.85) -- ( 55.99, 40.30);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 44.52, 62.12) --
	( 44.52, 51.85) --
	( 67.46, 51.85) --
	( 67.46, 62.12) --
	( 44.52, 62.12) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 44.52, 58.90) -- ( 67.46, 58.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48, 87.61) -- ( 81.48,119.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.48, 65.47) -- ( 81.48, 48.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 70.01, 87.61) --
	( 70.01, 65.47) --
	( 92.95, 65.47) --
	( 92.95, 87.61) --
	( 70.01, 87.61) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 70.01, 75.57) -- ( 92.95, 75.57);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 46.98) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 41.07) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 51.98) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 81.03) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 42.85) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 45.98) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 45.77) circle (  1.96);

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (123.96, 37.84) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 65.55) -- (123.96, 70.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.96, 60.24) -- (123.96, 56.21);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (112.49, 65.55) --
	(112.49, 60.24) --
	(135.43, 60.24) --
	(135.43, 65.55) --
	(112.49, 65.55) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (112.49, 62.49) -- (135.43, 62.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45, 96.67) -- (149.45,120.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.45, 80.54) -- (149.45, 67.34);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (137.98, 96.67) --
	(137.98, 80.54) --
	(160.92, 80.54) --
	(160.92, 96.67) --
	(137.98, 96.67) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.98, 91.07) -- (160.92, 91.07);
\definecolor{drawColor}{RGB}{51,51,51}
\definecolor{fillColor}{RGB}{51,51,51}

\path[draw=drawColor,draw opacity=0.70,line width= 0.4pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.70] (191.94, 85.35) circle (  1.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 69.95) -- (191.94, 78.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.94, 63.10) -- (191.94, 55.46);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (180.47, 69.95) --
	(180.47, 63.10) --
	(203.41, 63.10) --
	(203.41, 69.95) --
	(180.47, 69.95) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (180.47, 66.63) -- (203.41, 66.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43,107.50) -- (217.43,121.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.43, 92.91) -- (217.43, 72.38);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (205.96,107.50) --
	(205.96, 92.91) --
	(228.90, 92.91) --
	(228.90,107.50) --
	(205.96,107.50) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (205.96,101.27) -- (228.90,101.27);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 37.27) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 65.02) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90, 92.77) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 23.90,120.52) {3.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 68.73, 16.10) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (136.71, 16.10) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (204.68, 16.10) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (136.71,  6.25) {Signal-to-Noise Ratio};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70, 75.35) {Time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 74.54) --
	(266.20, 76.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 83.94) --
	(266.20, 86.10);
\definecolor{fillColor}{gray}{0.60}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 76.71) rectangle (271.62, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 80.32) --
	(271.62, 80.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 60.09) --
	(266.20, 62.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.20, 69.48) --
	(266.20, 71.65);
\definecolor{fillColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (260.78, 62.26) rectangle (271.62, 69.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.78, 65.87) --
	(271.62, 65.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 77.84) {Look-Ahead};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,130.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (277.92, 63.39) {Standard};
\end{scope}
\end{tikzpicture}

\end{document}