### 2021-06-24 paper/revision-comments.tex

\PassOptionsToPackage{unicode,pdfusetitle}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}

\documentclass[english,a4paper]{article}

\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm,mathtools,isomath}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{textcomp} % provide euro and other symbols
\usepackage{babel}

\usepackage{upquote} % straight quotes in verbatim environments
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts

\usepackage{xcolor}
\usepackage{xurl}
\usepackage{bookmark}
\usepackage{csquotes}

\usepackage{listings}

\lstset{
  basicstyle=\ttfamily, % print whole listing small
  %keywordstyle=\color{black}\bfseries\underbar, % underlined bold black keywords
  %commentstyle=\color{grey}, % white comments
  keepspaces = true % no special string spaces
}

\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue
}

\usepackage{longtable,booktabs}

\usepackage{etoolbox}

% nice tables out of csv files
\usepackage{csvsimple}

% format units nicely
\usepackage{siunitx}

% Allow footnotes in longtable head/foot
\usepackage{footnotehyper}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\usepackage{subcaption}

% tikz and pgfplots
\usepackage{pgfplots}
\pgfplotsset{width=7cm,compat=1.16}
\usepgfplotslibrary{external}
\tikzexternalize

\usepackage{pdfpages}

% algorithms
\usepackage{algorithm,algpseudocode}

% bibliography
\makeatletter
\def\blx@nowarnpolyglossia{}
\makeatother
\usepackage[style=alphabetic,backend=biber,sorting=anyt]{biblatex}
% \addbibresource{}

% \setlength{\emergencystretch}{3em} % prevent overfull lines

% nice author blocks
\usepackage[]{authblk}
\renewcommand\Affilfont{\itshape\small}

% title block
\title{Response to Referee}

\author{Johan Larsson}
\affil{The Department of Statistics, Lund University}

\date{\today}

\begin{document}

\maketitle

\noindent I would like to begin by thanking the referee for their valuable and insightful
comments on the paper. Here, I will outline all edits made to the paper based on
their comments, as well as any other changes to the paper.

\begin{displayquote}
  However, in Section 3, just one method (active warm start version of the Gap
  Safe rules) is used to compare the performance of the proposed strategy. The
  author should rephrase the last sentence of the abstract accordingly, e.g.:
  ``In experiments we show that these look-ahead screening rules outperform the
  active warm start version of the Gap Safe rules.''
\end{displayquote}
The last sentence of the abstract has been modified accordingly.

\begin{displayquote}
  Only the Gaussian case is considered. Is the method applicable to Generalized
  Linear Models?
\end{displayquote}
We have added a couple of sentences in the discussion discussing this.

\begin{displayquote}
  There are two options in the glmnet package for computing the path to Gaussian
  linear models: \texttt{type.gaussian="covariance"} and \texttt{type.gaussian="naive"}.
  Depending on the setting, these options have a large impact on timing. The
  latter can be far more efficient for p > n situations, or when p > 500. Are
  the results reported in simulations consistent under both settings?
\end{displayquote}
A sentence in the discussion has been added regarding this.

In addition to these changes, we have also replaced \(\tilde\beta\)
throughout the paper with \(\beta\), since the distinction made with the former
notation made no difference to the results.

We have moreover hade a few minor cosmetic changes to math environments,
figures, and text in order to maintain the page limit requirement. The
description of the experiment setup was also shortened somewhat for the same
reason.

\end{document}